---
title: "Record Linkage"
author: "Vaughn Johnson"
date: "4/30/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(here)
library(RecordLinkage)

path_to_src = here::here(file.path('R'))
load(file.path(path_to_src,
               "HarmonizedFiles",
               "HarmonizedDataSets.RData"))

common_cols = intersect(colnames(kbp_harmonized),
                        colnames(wapo_harmonized))
```

```{r}
common_cols = intersect(colnames(kbp_harmonized),
                        colnames(wapo_harmonized))

kbp_link  =  kbp_harmonized %>% select(common_cols) %>% select(-c("aka"))
wapo_link = wapo_harmonized %>% select(common_cols) %>% select(-c("aka"))
```

Here, we confirm there is at least 1 record linked between the two datasets
```{r}
### Confirm at least one person is present in both data sets
kbp_harmonized %>% filter(name == "Terry Lee Cockrell")
wapo_harmonized %>% filter(name == "Terry Lee Cockrell")
```

# Beginning the Record Linkage
Each record linkage process begins with a call to `compare.dedup` or 
`compare.linkage`. This returns an object of the class `RecLinkPairs`.

`RecLinkPairs` is literally undocumented in the RecordLinkage Manual. However,
`RecLinkData` is documented, and claims that this is the class returned by the
`compare.X` functions. So we'll assume we're working with `RecLinkData` objects.

`RecLinkData` have the following fields, though most of these are blank when you 
call `compare.linkage`, and are only filled in by future function calls.

|field|Description|Calculted by the function|
|---|---|---|
|`data1, data2`| Original Data Sets, unmutated. In order to even call `compare.linkage`, assume these have the same columns|`compare.linkage`|
|`pairs`|A `data.frame`. Each row is a pair of rows from `data1` and `data2`. Each column that `data1` and `data2` share has a number between 0 and 1. 1 means exactly the same, 0 means totally orthogonal|`compare.linkage`|
|`M`|Vector of $m$-probabilities.  $m$-probabilities are The m probability is the probability of agreement for a given matching variable when the comparison pair is a match|`emWeights`|
|`U`|Vector of $u$-probabilities. $u$-probabilities are The probability of this variable agreeing purely by chance for a comparison pair of two records not belonging to the same individual|`emWeights`|
|`W`| " Vector of log-likelihood weights as calculated by `emWeights`, corresponding to binary comparison patterns as created by bincombinations.|`emWeights`|
|`Wdata`|Vector of log-likelihood weights as calculated by `emWeights`. This is the sum of the weights of each column for each pair|`emWeights`|

`RecLinkData` objects are then passed to `emWeights`, which begins by doing the following

Each pairs has an $mu$ and a $u$. Basically, we guess $m$ and $u$ for each pair, and find the liklihood of 
our data based on those estiamtes. Then, we take the log of that likelihood, then the derivative,
and set it equal to 0. We then solve for the parameters again, and plug those back into a new liklihood
function. This process is repeated on the parameters converge. This whole process takes place
in `emWeights`.

Once we estimate as $k$ of the $m$ $u$ (two parameters for each column), we then
compute weights for each pair using something like log-odds

$$
w_{a,k} = \frac{\ln{(m_k / u_k)}}{\ln{(2)}}
$$

$$
w_{d,k} = \frac{\ln{((1 - m_k) / (1 - u_k))}}{\ln{(2)}}
$$

Where $a$ means agree, and $d$ means disagree. Then we sum alllll the parameters
together for each pair, which is put in `Wdata`. The more positive, the the more 
likely a match.

```{r}
blank_linkage_object = compare.linkage(kbp_link, 
                                 wapo_link, 
                                 blockfld = c('state'), 
                                 phonetic = T,
                                 strcmp =T)

EM_linkage_object = emWeights(blank_linkage_object)
em_weights = EM_linkage_object$Wdata
name_match = EM_linkage_object$pairs$name == 1
weight_names = data.frame(em_weights, name_match)

ggplot(weight_names, aes(x=em_weights, group=name_match, fill=name_match)) + 
    geom_histogram(bins=50) +
    scale_y_log10() +
    ggtitle("Histogram of Weights assigned to record pairs, log scale")
```

## Clerical Review

```{r}

record_pairs = EM_linkage_object$pairs %>% select(c('id1', 'id2'))

record_pairs$weight = EM_linkage_object$Wdata

record_pairs$name_1 = EM_linkage_object$data1[record_pairs$id1, 'name']
record_pairs$name_2 = EM_linkage_object$data1[record_pairs$id2, 'name']


record_pairs %>% filter(weight > 0 & weight < 20) %>% head()
```





# Bibliography
Improving EM Algorithm Estimates for Record Linkage Parameters (2002)
- EM stuff on page 3

Probabilistic record linkage and a method to calculate the positive predictive value
- $M$ and $U$ explanations in "Record linkage methodology" section

Analysis of a Probabilistic Record Linkage Technique without Human Review
- Parameter fitting in EM in "METHODS" section



