---
title: "Fatal Encounters with Police"
author: "Madeline Cummins"
output:
  xaringan::moon_reader:
    yolo: FALSE     
    css: UGRSMaddi.css  # our .css file
    lib_dir: libs
---
setwd("F:/GitHub/FFSG/ffsg")

class: left, top
### Our Goals
.left-column[
Investigate and understand how different locations and demographics are affected by police violence

Not much data available - crowd sourced
]

---

class: left, top
### Police Related Deaths Per Capita in Washington
.left-column[
More per capita than the US average

Initiative 940
]

.right-column[
```{r, echo=FALSE}
source('Analysis/Tables/permillcalculation.R')
source('Analysis/Graphics/permillgraphfunc.R')
permillgraph("Washington", all = TRUE, capita = TRUE)
```
]

---

class: left, top
### Race and Gender
.pull-left[
Race

```{r, echo=FALSE}
source(here('Analysis/Graphics/descstatfuncs.R'))
dsplot("Race")
```
]

.pull-right[
Gender

```{r}
source(here('Analysis/Graphics/descstatfuncs.R'))
dsplot("Gender")
```
]

---

class: left, top
### Choropleth
.left-column[
Shade of state is determined by per capita values for each state

Shows for specific years or the mean
]

.right-column[
```{r}
source(here('Analysis/Graphics/choroplethmapfunc.R'))
choroplethmap()
```
]

---

class: left, top
### Cartogram
.left-column[
Size of state is determined by per capita values for each state

Shows for specific years or the mean

Harder to create but gives a much more intense visual
]

.right-column[
```{r}
source('Analysis/Graphics/cartogramfunc.R')
ffcartogram(2017)
```
]

---

class: left, top
### Interactive Map
.column[
```{r}
library(leaflet)
source('Analysis/Graphics/interactivemap.R')
interactivemap
```
]

---

class: left, top
### Future Work
.column[
Deal with missing data

Better cartogram

More analysis
]

---

class: left, top
### Jainul's Work
