---
title: "Fatal Encounters with Police"
author: "Madeline Cummins"
output:
  xaringan::moon_reader:
    yolo: FALSE     
    lib_dir: libs
    
---

#I am trying to get rid of this slide but am unsure of how to make the black US not render
```{r setup, echo = FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(ggplot2)
library(here)
library(dplyr)
opts_chunk$set(message = FALSE,
               warning = FALSE)
```

```{r, echo=FALSE}
source(here('Analysis/Tables/permillcalculation.R'))
```

```{r, echo=FALSE, results='hide', message=FALSE}
source(here('Analysis/Graphics/cartogramfunc.R'))
```
---
class: left, top
### Outline
.column[
Investigate and understand how different locations and demographics are affected by police violence

Make data easily available to public

Shiny app with tables, static visualtions, and interactive maps to display the data
]

---

class: left, top
### Tabular Representation
.pull-left[
```{r, echo=FALSE}
source(here('Analysis/Tables/permilltablefunc.R'))
permilltable("Washington", TRUE, FALSE)
#I will shorten this so it is a more consise ~ 2-3 columns
```
]

.pull-right[
```{r, echo=FALSE}
source(here('Analysis/Graphics/descstatfuncs.R'))
dstable("Race")
```
]

---

class:left, top
### Static Plots
.pull-left[
```{r, echo=FALSE}
source(here('Analysis/Graphics/permillgraphfunc.R'))
permillgraph("United States", FALSE, TRUE)
```
]

.pull-right[
```{r, echo=FALSE}
source(here('Analysis/Graphics/permillgraphfunc.R'))
permillgraph("Washington", TRUE, TRUE)
```
]

---

class: left, top
### Choropleth
.column[
```{r, echo=FALSE}
source(here('Analysis/Graphics/choroplethmapfunc.R'))
choroplethmap()
```
]
---

class: left, top
### Cartogram
.column[
```{r, echo=FALSE}
cart2016
```
]
---
### Present Shiny Here
---
class:left, top
### Future Work
.column[
Make work publicaly available

Compile into a package with guide

Facebook page

Linked with fatal encounters, killed by police, and campaign 0

Add pics to interactive map
]