---
title: "linker Report"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r, include= FALSE}
library(DataExplorer)
library(networkD3)
library(dplyr)
library(rlist)
library(data.table)
library(DT)
library(plotly)

path_to_src = here::here(file.path( 'Data'))
load(file=file.path(path_to_src,
                    "FinalClassification",
                    "full_classification.RData"))

load(file=file.path(path_to_src,
                    "FinalClassification",
                    "full_combined_harmonized.RData"))

load(file= file.path(path_to_src,
                     "HarmonizedFiles",
                     "HarmonizedDataSets.RData"))
```


## Overview

In this document, we will go through the outputs generated from `Linker.R`.
When running `Linker.R`, we get two outputs: classification and combined_harmonized.
for a more detailed description of how each variable is calculated using RecordLinkage, please look over these resources: [resource 1](https://cran.r-project.org/web/packages/RecordLinkage/RecordLinkage.pdf), [resource 2](http://v-neck.github.io/final/),
[resource 3](https://docs.google.com/presentation/d/19XFzorND9YIxosJm74sgS-We_qJSvCoGWwsGXubt-N0/edit#slide=id.p)


### Combined harmonized outputs:

before going through combined_harmonized output, let's first see the common columns to all 4 datasets:

Common columns :
```{r}
Reduce(intersect, list(colnames(kbp_harmonized),
                                     colnames(fe_harmonized),
                                     colnames(mpv_harmonized),
                                     colnames(wapo_harmonized)))
```



`Combined harmonized`is a data frame in which each of the four harmonized datasets is stacked row-wise.


```{r}
str <- plot_str(combined_harmonized,
                print_network = FALSE)
diagonalNetwork(str)

```

**Breakout of common columns in combined_harmonized dataset**

```{r}

datatable(dcast(combined_harmonized,
                  sex ~source,
                  value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)

datatable(dcast(combined_harmonized,
                  race ~source,
                  value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)

datatable(dcast(combined_harmonized,
                  state ~ source,
                  value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)

datatable(dcast(combined_harmonized,
                  month~ source,
                  value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)

datatable(dcast(combined_harmonized,
                  day ~ source,
                  value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)

datatable(dcast(combined_harmonized,
                   year~ source,
                   value.var = 'uid',
                  fun.aggregate = length),
          rownames = FALSE)
```






### Classification outputs:

Classification is a list of 10 components, below is the components of this list:
```{r}
list.names(classification)
```

* *Data*: is the original data that we want to perform record linkage on
```{r}
str <- plot_str(classification$data,
                print_network = FALSE)
diagonalNetwork(str)
```

* *Pairs*: is a dataframe of all possible pairs in _data_.
Each row represents the comparison pattern of two records, identified by columns *id1* and *id2*. The other columns contain for each considered attribute a real number in the range [0..1] representing the degree of similarity. The last column
contains the matching status of the pair, coded as 1 for a match or 0 for a non-match.
```{r}
str <- plot_str(classification$pairs,
                print_network = FALSE)
diagonalNetwork(str)

```


* *Frequencies*: the average frequency of values for each column included in pairs.
```{r}
classification$frequencies
```

* *type*: to identify which process is done: 'deduplication' or 'linkage'
```{r}
classification$type
```

* *M*: the conditional probability of a match given the observed difference in attributes.
```{r}
par(mfrow=c(1,2))
boxplot(log(classification$M))
plot(density(log(classification$M)))

```

* *U*: the probabilty given the pair is a nonmatch. 
As the values were skewed, we had to log the y-axis.
```{r}
par(mfrow=c(1,2))
boxplot(log(classification$U))
plot(density(log(classification$U)))

```



* *M vs U*: 
```{r}
plot(log(classification$M),log(classification$U))
```

* *W*: is the weight for each pair. it is calculated using the function emWeights in the package recordLinkage.
W is the log of M divided by U.

```{r}
par(mfrow=c(1,2))
boxplot(classification$W)
plot(density(classification$W))

hist(classification$W, breaks = 400 )
```

* *Wdata*: Vector of log-likelihood weights as calculated by emWeights,
corresponding to the rows of pairs.

```{r}
par(mfrow=c(1,2))
plot(density(classification$Wdata))
hist(classification$Wdata)
```


* *Prediction*: Linkage result. Coded as a factor with levels "N" for non-links, "P" for possible links and "L" for links.

```{r}
table(classification$prediction)
```


**Pairs between -21 and -19**
```{r}
classification$pairs$weight <- classification$Wdata

filtered <- classification$pairs %>% 
            filter(weight >= -21, weight <= -20) %>% 
            select(id1,
                   id2)


cb <- combined_harmonized %>%
  select(firstname,
         lastname,
         sex,
         age,
         state,
         race,
         day,
         year,
         source,
         uid)

join <-  inner_join(filtered,
                    cb,
                    by= c("id1" = "uid"))

second_join <- inner_join( join,
                           cb,
                           by= c("id2" = "uid"),
                           suffix = c(".1", ".2"))


datatable(dcast(second_join,
      source.1 ~source.2,
      fun.aggregate = length,
      value.var = "sex.2"),
      rownames = FALSE)

datatable(dcast(second_join,
      sex.1 ~sex.2,
      fun.aggregate = length,
      value.var = "sex.2"),
      rownames = FALSE)

datatable(dcast(second_join,
      race.1 ~race.2,
      fun.aggregate = length,
      value.var = "state.2"),
      rownames = FALSE)

datatable(dcast(second_join,
      state.1 ~state.2,
      fun.aggregate = length,
      value.var = "state.2"),
      rownames = FALSE)

datatable(dcast(second_join,
      year.1 ~year.2,
      fun.aggregate = length,
      value.var = "year.2"),
      rownames = FALSE)


randomSample <- second_join %>% 
  
  sample_n(10000)

ggplot(randomSample , aes(x = age.1 ,y =age.2)) +
  geom_point() 



plot_ly(data = second_join,
        x = ~age.1,
        y = ~age.2)

```




